---
title: "Benchmarking ML Performance: Original vs Anonymized Data"
output: html_notebook
---

This notebook compares the predictive performance of ML models on a variety of
datasets, before and after anonymization with safetrain's cryptoencoder method.
It covers classification and regression tasks across different dataset sizes,
column counts, and data characteristics.

## Setup

```{r setup, message=FALSE, warning=FALSE}
library(safetrain)
library(mlbench)
library(ranger)
library(gbm)
library(glmnet)
library(MASS)

set.seed(42)
key <- generate_key()

N_CV_FOLDS <- 5
N_REPEATS  <- 3
```

## Helper Functions

```{r helpers}
#' Prepare a dataset: returns list with data, target column name, and task type.
prepare_dataset <- function(name) {

  if (name == "BreastCancer") {
    data(BreastCancer, package = "mlbench", envir = environment())
    df <- BreastCancer
    df$Id <- NULL
    for (col in setdiff(names(df), "Class")) {
      df[[col]] <- as.numeric(as.character(df[[col]]))
    }
    df <- na.omit(df)
    list(data = df, target = "Class", task = "classification")

  } else if (name == "PimaIndiansDiabetes") {
    data(PimaIndiansDiabetes, package = "mlbench", envir = environment())
    df <- PimaIndiansDiabetes
    list(data = df, target = "diabetes", task = "classification")

  } else if (name == "BostonHousing") {
    data(BostonHousing, package = "mlbench", envir = environment())
    df <- BostonHousing
    df$chas <- as.numeric(as.character(df$chas))
    list(data = df, target = "medv", task = "regression")

  } else if (name == "Ionosphere") {
    data(Ionosphere, package = "mlbench", envir = environment())
    df <- Ionosphere
    # V2 is constant zero in this dataset
    df$V2 <- NULL
    for (col in setdiff(names(df), "Class")) {
      df[[col]] <- as.numeric(df[[col]])
    }
    list(data = df, target = "Class", task = "classification")

  } else if (name == "Sonar") {
    data(Sonar, package = "mlbench", envir = environment())
    df <- Sonar
    list(data = df, target = "Class", task = "classification")

  } else if (name == "Glass") {
    data(Glass, package = "mlbench", envir = environment())
    df <- Glass
    # Binary: window vs non-window glass
    df$Type <- factor(ifelse(df$Type %in% c(1, 2, 3), "window", "non-window"))
    list(data = df, target = "Type", task = "classification")

  } else if (name == "Vehicle") {
    data(Vehicle, package = "mlbench", envir = environment())
    df <- Vehicle
    # Binary: bus vs non-bus
    df$Class <- factor(ifelse(df$Class == "bus", "bus", "other"))
    list(data = df, target = "Class", task = "classification")

  } else if (name == "Servo") {
    data(Servo, package = "mlbench", envir = environment())
    df <- Servo
    for (col in names(df)) {
      if (is.factor(df[[col]]) && col != "class") {
        df[[col]] <- as.numeric(df[[col]])
      }
    }
    names(df)[names(df) == "class"] <- "y"
    list(data = df, target = "y", task = "regression")

  } else if (name == "Ozone") {
    data(Ozone, package = "mlbench", envir = environment())
    df <- Ozone
    # Convert all factors to numeric
    for (col in names(df)) {
      if (is.factor(df[[col]])) df[[col]] <- as.numeric(df[[col]])
    }
    df <- na.omit(df)
    names(df)[1] <- "ozone"
    list(data = df, target = "ozone", task = "regression")

  } else if (name == "Synthetic_Wide") {
    # Synthetic dataset: 2000 rows, 50 numeric features
    n <- 2000; p <- 50
    X <- matrix(rnorm(n * p), nrow = n)
    colnames(X) <- paste0("V", seq_len(p))
    beta <- rnorm(p) * c(rep(1, 10), rep(0.1, p - 10))
    y <- X %*% beta + rnorm(n, sd = 2)
    df <- as.data.frame(X)
    df$y <- as.numeric(y)
    list(data = df, target = "y", task = "regression")

  } else if (name == "Synthetic_Large") {
    # Synthetic dataset: 10000 rows, 20 features, binary classification
    n <- 10000; p <- 20
    X <- matrix(rnorm(n * p), nrow = n)
    colnames(X) <- paste0("V", seq_len(p))
    logit <- X[, 1] * 0.8 + X[, 2] * 0.6 - X[, 3] * 0.5 +
             X[, 1] * X[, 2] * 0.3 + rnorm(n, sd = 0.5)
    prob <- 1 / (1 + exp(-logit))
    y <- factor(ifelse(rbinom(n, 1, prob) == 1, "pos", "neg"))
    df <- as.data.frame(X)
    df$y <- y
    list(data = df, target = "y", task = "classification")

  } else if (name == "Synthetic_VeryWide") {
    # Synthetic dataset: 1000 rows, 100 features
    n <- 1000; p <- 100
    X <- matrix(rnorm(n * p), nrow = n)
    colnames(X) <- paste0("V", seq_len(p))
    beta <- rep(0, p)
    beta[1:15] <- rnorm(15)
    y <- X %*% beta + rnorm(n, sd = 3)
    df <- as.data.frame(X)
    df$y <- as.numeric(y)
    list(data = df, target = "y", task = "regression")

  } else if (name == "Synthetic_Huge") {
    # Large classification: 50000 rows, 30 features
    n <- 50000; p <- 30
    X <- matrix(rnorm(n * p), nrow = n)
    colnames(X) <- paste0("V", seq_len(p))
    logit <- X[, 1] * 1.0 + X[, 2] * 0.7 - X[, 3] * 0.5 +
             X[, 4] * X[, 5] * 0.4 + sin(X[, 6]) * 0.6 + rnorm(n, sd = 0.5)
    prob <- 1 / (1 + exp(-logit))
    y <- factor(ifelse(rbinom(n, 1, prob) == 1, "pos", "neg"))
    df <- as.data.frame(X)
    df$y <- y
    list(data = df, target = "y", task = "classification")

  } else if (name == "Synthetic_VeryLarge_Reg") {
    # Large regression: 20000 rows, 40 features
    n <- 20000; p <- 40
    X <- matrix(rnorm(n * p), nrow = n)
    colnames(X) <- paste0("V", seq_len(p))
    beta <- c(rnorm(10, sd = 2), rnorm(10, sd = 0.5), rep(0, p - 20))
    y <- X %*% beta + X[, 1] * X[, 2] * 0.5 + rnorm(n, sd = 2)
    df <- as.data.frame(X)
    df$y <- as.numeric(y)
    list(data = df, target = "y", task = "regression")

  } else if (name == "Synthetic_UltraWide") {
    # p >> n scenario: 500 rows, 200 features
    n <- 500; p <- 200
    X <- matrix(rnorm(n * p), nrow = n)
    colnames(X) <- paste0("V", seq_len(p))
    beta <- rep(0, p)
    beta[1:8] <- rnorm(8, sd = 2)
    y <- X %*% beta + rnorm(n, sd = 2)
    df <- as.data.frame(X)
    df$y <- as.numeric(y)
    list(data = df, target = "y", task = "regression")

  } else if (name == "Synthetic_HighDim_Clf") {
    # High-dimensional classification: 5000 rows, 100 features
    n <- 5000; p <- 100
    X <- matrix(rnorm(n * p), nrow = n)
    colnames(X) <- paste0("V", seq_len(p))
    logit <- X[, 1:5] %*% c(1.2, 0.8, -0.6, 0.4, -0.3) +
             X[, 1] * X[, 3] * 0.5 + rnorm(n, sd = 0.7)
    prob <- 1 / (1 + exp(-logit))
    y <- factor(ifelse(rbinom(n, 1, prob) == 1, "pos", "neg"))
    df <- as.data.frame(X)
    df$y <- y
    list(data = df, target = "y", task = "classification")

  } else if (name == "Synthetic_Correlated") {
    # Correlated features: 3000 rows, 30 features with block correlation
    n <- 3000; p <- 30
    sigma <- diag(p)
    # Create 3 blocks of 10 correlated features
    for (block in 0:2) {
      idx <- (block * 10 + 1):(block * 10 + 10)
      sigma[idx, idx] <- 0.6
      diag(sigma[idx, idx]) <- 1
    }
    X <- MASS::mvrnorm(n, mu = rep(0, p), Sigma = sigma)
    colnames(X) <- paste0("V", seq_len(p))
    beta <- c(rnorm(10, sd = 1.5), rnorm(10, sd = 0.5), rep(0, 10))
    y <- X %*% beta + rnorm(n, sd = 2)
    df <- as.data.frame(X)
    df$y <- as.numeric(y)
    list(data = df, target = "y", task = "regression")

  } else if (name == "Synthetic_Sparse_Signal") {
    # Sparse signal: 4000 rows, 80 features, only 5 informative
    n <- 4000; p <- 80
    X <- matrix(rnorm(n * p), nrow = n)
    colnames(X) <- paste0("V", seq_len(p))
    logit <- X[, 1] * 1.5 + X[, 2] * 1.0 + X[, 3] * 0.8 +
             X[, 4] * 0.6 + X[, 5] * 0.4 + rnorm(n, sd = 0.5)
    prob <- 1 / (1 + exp(-logit))
    y <- factor(ifelse(rbinom(n, 1, prob) == 1, "pos", "neg"))
    df <- as.data.frame(X)
    df$y <- y
    list(data = df, target = "y", task = "classification")

  } else if (name == "Synthetic_Nonlinear") {
    # Nonlinear regression: 3000 rows, 20 features
    n <- 3000; p <- 20
    X <- matrix(rnorm(n * p), nrow = n)
    colnames(X) <- paste0("V", seq_len(p))
    y <- sin(X[, 1] * 2) + X[, 2]^2 + exp(X[, 3] * 0.5) +
         X[, 4] * X[, 5] + log(abs(X[, 6]) + 1) + rnorm(n, sd = 0.5)
    df <- as.data.frame(X)
    df$y <- as.numeric(y)
    list(data = df, target = "y", task = "regression")

  } else if (name == "Synthetic_Imbalanced") {
    # Imbalanced classification: 5000 rows, 25 features, 10% positive
    n <- 5000; p <- 25
    X <- matrix(rnorm(n * p), nrow = n)
    colnames(X) <- paste0("V", seq_len(p))
    logit <- X[, 1] * 1.0 + X[, 2] * 0.8 - X[, 3] * 0.6 - 2.0  # shift for imbalance
    prob <- 1 / (1 + exp(-logit))
    y <- factor(ifelse(rbinom(n, 1, prob) == 1, "pos", "neg"))
    df <- as.data.frame(X)
    df$y <- y
    list(data = df, target = "y", task = "classification")
  }
}

#' Anonymize training features and return both data and mapping.
anonymize_train <- function(features, key) {
  anonymize_data(features, key, opts = list(method = "cryptoencoder"))
}

#' Apply an existing mapping to test features (no retraining).
anonymize_test <- function(features, key, mapping) {
  transform_new_data(features, key, mapping)
}

#' Compute classification metrics.
classification_metrics <- function(preds, actual, model_obj = NULL,
                                   test_data = NULL, target = NULL,
                                   task_formula = NULL) {
  acc <- mean(preds == actual)
  lvls <- levels(actual)

  # Confusion matrix for binary tasks
  if (length(lvls) == 2) {
    tp <- sum(preds == lvls[2] & actual == lvls[2])
    fp <- sum(preds == lvls[2] & actual == lvls[1])
    fn <- sum(preds == lvls[1] & actual == lvls[2])
    tn <- sum(preds == lvls[1] & actual == lvls[1])

    precision <- if ((tp + fp) > 0) tp / (tp + fp) else 0
    recall    <- if ((tp + fn) > 0) tp / (tp + fn) else 0
    f1        <- if ((precision + recall) > 0) {
      2 * precision * recall / (precision + recall)
    } else {
      0
    }

    # Balanced accuracy
    spec <- if ((tn + fp) > 0) tn / (tn + fp) else 0
    sens <- recall
    bal_acc <- (sens + spec) / 2

    # Matthews Correlation Coefficient
    mcc_num <- (tp * tn) - (fp * fn)
    mcc_den <- sqrt((tp + fp) * (tp + fn) * (tn + fp) * (tn + fn))
    mcc <- if (mcc_den > 0) mcc_num / mcc_den else 0

    # Cohen's Kappa
    n_total <- tp + fp + fn + tn
    p_obs <- (tp + tn) / n_total
    p_yes <- ((tp + fp) / n_total) * ((tp + fn) / n_total)
    p_no  <- ((fn + tn) / n_total) * ((fp + tn) / n_total)
    p_exp <- p_yes + p_no
    kappa <- if (p_exp < 1) (p_obs - p_exp) / (1 - p_exp) else 0

    c(Accuracy = acc, Balanced_Acc = bal_acc,
      Precision = precision, Recall = recall, F1 = f1,
      MCC = mcc, Kappa = kappa)
  } else {
    c(Accuracy = acc, Balanced_Acc = NA,
      Precision = NA, Recall = NA, F1 = NA,
      MCC = NA, Kappa = NA)
  }
}

#' Compute regression metrics.
regression_metrics <- function(preds, actual) {
  residuals <- actual - preds
  n <- length(actual)
  rmse <- sqrt(mean(residuals^2))
  mae  <- mean(abs(residuals))
  medae <- median(abs(residuals))
  max_err <- max(abs(residuals))
  ss_res <- sum(residuals^2)
  ss_tot <- sum((actual - mean(actual))^2)
  r2 <- 1 - ss_res / ss_tot
  # MAPE, guarding against zero actuals
  nonzero <- actual != 0
  mape <- if (any(nonzero)) mean(abs(residuals[nonzero] / actual[nonzero])) * 100 else NA
  # Explained variance score
  ev <- 1 - var(residuals) / var(actual)

  c(RMSE = rmse, MAE = mae, MedAE = medae, Max_Error = max_err,
    R2 = r2, MAPE = mape, Expl_Var = ev)
}

#' Fit three models and return a data frame of performance metrics.
fit_models <- function(train, test, target, task) {
  formula <- as.formula(paste(target, "~ ."))
  all_metrics <- list()

  # --- ranger (random forest) ---
  if (task == "classification") {
    rf <- ranger(formula, data = train, num.trees = 500,
                 classification = TRUE, seed = 1)
    preds_rf <- predict(rf, data = test)$predictions
    all_metrics[["ranger"]] <- classification_metrics(preds_rf, test[[target]])
  } else {
    rf <- ranger(formula, data = train, num.trees = 500, seed = 1)
    preds_rf <- predict(rf, data = test)$predictions
    all_metrics[["ranger"]] <- regression_metrics(preds_rf, test[[target]])
  }

  # --- gbm (gradient boosting) ---
  if (task == "classification") {
    lvls <- levels(train[[target]])
    train_gbm <- train
    train_gbm[[target]] <- as.integer(train_gbm[[target]]) - 1L
    gb <- gbm(formula, data = train_gbm, distribution = "bernoulli",
              n.trees = 500, interaction.depth = 3, shrinkage = 0.05,
              n.minobsinnode = 10, verbose = FALSE)
    prob <- predict(gb, newdata = test, n.trees = 500, type = "response")
    preds_gb <- factor(lvls[(prob > 0.5) + 1], levels = lvls)
    all_metrics[["gbm"]] <- classification_metrics(preds_gb, test[[target]])
  } else {
    gb <- gbm(formula, data = train, distribution = "gaussian",
              n.trees = 500, interaction.depth = 3, shrinkage = 0.05,
              n.minobsinnode = 10, verbose = FALSE)
    preds_gb <- predict(gb, newdata = test, n.trees = 500)
    all_metrics[["gbm"]] <- regression_metrics(preds_gb, test[[target]])
  }

  # --- glmnet (elastic net) ---
  x_train <- model.matrix(formula, data = train)[, -1, drop = FALSE]
  x_test  <- model.matrix(formula, data = test)[, -1, drop = FALSE]
  if (task == "classification") {
    y_train <- train[[target]]
    gl <- cv.glmnet(x_train, y_train, family = "binomial", alpha = 0.5)
    prob_gl <- predict(gl, newx = x_test, s = "lambda.min", type = "response")
    lvls <- levels(y_train)
    preds_gl <- factor(lvls[(prob_gl > 0.5) + 1], levels = lvls)
    all_metrics[["glmnet"]] <- classification_metrics(preds_gl, test[[target]])
  } else {
    y_train <- train[[target]]
    gl <- cv.glmnet(x_train, y_train, family = "gaussian", alpha = 0.5)
    preds_gl <- as.numeric(predict(gl, newx = x_test, s = "lambda.min"))
    all_metrics[["glmnet"]] <- regression_metrics(preds_gl, test[[target]])
  }

  # --- LDA (classification only) / Ridge regression ---
  if (task == "classification") {
    lda_fit <- tryCatch({
      MASS::lda(formula, data = train)
    }, error = function(e) NULL)
    if (!is.null(lda_fit)) {
      preds_lda <- predict(lda_fit, newdata = test)$class
      all_metrics[["lda"]] <- classification_metrics(preds_lda, test[[target]])
    }
  } else {
    gl_ridge <- cv.glmnet(x_train, y_train, family = "gaussian", alpha = 0)
    preds_ridge <- as.numeric(predict(gl_ridge, newx = x_test, s = "lambda.min"))
    all_metrics[["ridge"]] <- regression_metrics(preds_ridge, test[[target]])
  }

  all_metrics
}
```

## Run Benchmark

We use stratified (classification) or random (regression) train/test splits.
Anonymization is applied separately to train and test to prevent leakage.

```{r benchmark, message=FALSE, warning=FALSE}
dataset_names <- c(
  # Real-world classification
  "BreastCancer", "PimaIndiansDiabetes", "Ionosphere", "Sonar",
  "Glass", "Vehicle",
  # Real-world regression
  "BostonHousing", "Ozone"
  # # Synthetic datasets (various sizes and characteristics)
  # "Synthetic_Wide", 
  # #"Synthetic_Large", 
  # "Synthetic_VeryWide"
  # #"Synthetic_Huge", 
  # # "Synthetic_VeryLarge_Reg", "Synthetic_UltraWide",
  # # #"Synthetic_HighDim_Clf", 
  # # "Synthetic_Correlated", "Synthetic_Sparse_Signal",
  # # "Synthetic_Nonlinear", "Synthetic_Imbalanced"
)

results <- list()
timing  <- list()

for (ds_name in dataset_names) {
  cat("===", ds_name, "===\n")
  ds <- prepare_dataset(ds_name)
  rownames(ds$data) <- NULL

  cat("  Rows:", nrow(ds$data), " Cols:", ncol(ds$data) - 1,
      " Task:", ds$task, "\n")

  y <- ds$data[[ds$target]]
  X <- ds$data[, setdiff(names(ds$data), ds$target), drop = FALSE]

  # Train/test split
  n <- nrow(X)
  if (ds$task == "classification") {
    train_idx <- unlist(tapply(seq_len(n), y, function(i) {
      sample(i, size = round(length(i) * 0.7))
    }))
  } else {
    train_idx <- sample(seq_len(n), size = round(n * 0.7))
  }
  test_idx <- setdiff(seq_len(n), train_idx)

  X_train <- X[train_idx, , drop = FALSE]
  X_test  <- X[test_idx, , drop = FALSE]
  y_train <- y[train_idx]
  y_test  <- y[test_idx]

  # Train the autoencoder on X_train only (target is in y_train, never seen
  # by the encoder). Then apply the same trained weights to X_test via the
  # mapping, so both sets share an identical representation space.
  t0 <- proc.time()
  res_train  <- anonymize_train(X_train, key)
  anon_train <- res_train$data
  anon_test  <- anonymize_test(X_test, key, res_train$mapping)
  t_anon <- (proc.time() - t0)[["elapsed"]]

  timing[[ds_name]] <- data.frame(
    Dataset = ds_name,
    Rows = nrow(ds$data),
    Features = ncol(X),
    Anon_Time_Sec = round(t_anon, 3),
    stringsAsFactors = FALSE
  )

  # Reassemble with target

  orig_train <- X_train; orig_train[[ds$target]] <- y_train
  orig_test  <- X_test;  orig_test[[ds$target]]  <- y_test
  anon_train[[ds$target]] <- y_train
  anon_test[[ds$target]]  <- y_test

  metrics_orig <- fit_models(orig_train, orig_test, ds$target, ds$task)
  metrics_anon <- fit_models(anon_train, anon_test, ds$target, ds$task)

  for (model_name in names(metrics_orig)) {
    mo <- metrics_orig[[model_name]]
    ma <- metrics_anon[[model_name]]
    metric_names <- names(mo)

    for (mn in metric_names) {
      results[[paste(ds_name, model_name, mn, sep = "|")]] <- data.frame(
        Dataset    = ds_name,
        Task       = ds$task,
        Features   = ncol(X),
        Rows       = nrow(ds$data),
        Model      = model_name,
        Metric     = mn,
        Original   = round(mo[[mn]], 4),
        Anonymized = round(ma[[mn]], 4),
        stringsAsFactors = FALSE
      )
    }
  }

  cat("  Anonymization time:", round(t_anon, 3), "sec\n\n")
}

all_results <- do.call(rbind, results)
rownames(all_results) <- NULL
all_timing <- do.call(rbind, timing)
rownames(all_timing) <- NULL
```

## Results: Classification Datasets

```{r results_classification}
clf_results <- all_results[all_results$Task == "classification", ]
knitr::kable(
  clf_results[, c("Dataset", "Rows", "Features", "Model", "Metric",
                   "Original", "Anonymized")],
  caption = "Classification: Original vs Anonymized"
)
```

## Results: Regression Datasets

```{r results_regression}
reg_results <- all_results[all_results$Task == "regression", ]
knitr::kable(
  reg_results[, c("Dataset", "Rows", "Features", "Model", "Metric",
                   "Original", "Anonymized")],
  caption = "Regression: Original vs Anonymized"
)
```

## Anonymization Timing

```{r timing_table}
knitr::kable(all_timing, caption = "Anonymization Time by Dataset Size")
```

## Metric Degradation Summary

How much does anonymization degrade each metric on average?

```{r degradation}
# For each metric, compute average absolute change across datasets and models
clf_acc <- clf_results[clf_results$Metric == "Accuracy", ]
clf_f1  <- clf_results[clf_results$Metric == "F1" & !is.na(clf_results$Original), ]
reg_rmse <- reg_results[reg_results$Metric == "RMSE", ]
reg_r2   <- reg_results[reg_results$Metric == "R2", ]

summary_df <- data.frame(
  Metric = c("Accuracy", "F1", "RMSE", "R-squared"),
  Avg_Original = c(
    mean(clf_acc$Original, na.rm = TRUE),
    mean(clf_f1$Original, na.rm = TRUE),
    mean(reg_rmse$Original, na.rm = TRUE),
    mean(reg_r2$Original, na.rm = TRUE)
  ),
  Avg_Anonymized = c(
    mean(clf_acc$Anonymized, na.rm = TRUE),
    mean(clf_f1$Anonymized, na.rm = TRUE),
    mean(reg_rmse$Anonymized, na.rm = TRUE),
    mean(reg_r2$Anonymized, na.rm = TRUE)
  ),
  stringsAsFactors = FALSE
)
summary_df$Avg_Original   <- round(summary_df$Avg_Original, 4)
summary_df$Avg_Anonymized <- round(summary_df$Avg_Anonymized, 4)
summary_df$Change <- round(summary_df$Avg_Anonymized - summary_df$Avg_Original, 4)

knitr::kable(summary_df, caption = "Average Metric Degradation from Anonymization")
```

## Visualization: Accuracy and F1 by Dataset

```{r plot_classification, fig.width=12, fig.height=8}
par(mfrow = c(2, 3), mar = c(6, 4, 3, 1))

clf_datasets <- unique(clf_results$Dataset)

for (ds_name in clf_datasets) {
  df_acc <- clf_results[clf_results$Dataset == ds_name &
                        clf_results$Metric == "Accuracy", ]
  bar_vals <- rbind(df_acc$Original, df_acc$Anonymized)
  colnames(bar_vals) <- df_acc$Model

  y_range <- range(bar_vals, na.rm = TRUE)
  y_pad   <- diff(y_range) * 0.15
  ylim    <- c(max(0, y_range[1] - y_pad), min(1, y_range[2] + y_pad))

  barplot(bar_vals, beside = TRUE, col = c("steelblue", "coral"),
          main = paste(ds_name, "- Accuracy"), ylab = "Accuracy", ylim = ylim,
          names.arg = df_acc$Model, las = 2,
          legend.text = c("Original", "Anonymized"),
          args.legend = list(x = "topright", cex = 0.7))
}
```

```{r plot_f1, fig.width=12, fig.height=8}
par(mfrow = c(2, 3), mar = c(6, 4, 3, 1))

for (ds_name in clf_datasets) {
  df_f1 <- clf_results[clf_results$Dataset == ds_name &
                       clf_results$Metric == "F1" &
                       !is.na(clf_results$Original), ]
  if (nrow(df_f1) == 0) next

  bar_vals <- rbind(df_f1$Original, df_f1$Anonymized)
  colnames(bar_vals) <- df_f1$Model

  y_range <- range(bar_vals, na.rm = TRUE)
  y_pad   <- diff(y_range) * 0.15
  ylim    <- c(max(0, y_range[1] - y_pad), min(1, y_range[2] + y_pad))

  barplot(bar_vals, beside = TRUE, col = c("steelblue", "coral"),
          main = paste(ds_name, "- F1 Score"), ylab = "F1", ylim = ylim,
          names.arg = df_f1$Model, las = 2,
          legend.text = c("Original", "Anonymized"),
          args.legend = list(x = "topright", cex = 0.7))
}
```

## Visualization: Regression Metrics

```{r plot_regression, fig.width=12, fig.height=10}
reg_datasets <- unique(reg_results$Dataset)
n_reg <- length(reg_datasets)

par(mfrow = c(n_reg, 2), mar = c(6, 4, 3, 1))

for (ds_name in reg_datasets) {
  # RMSE
  df_rmse <- reg_results[reg_results$Dataset == ds_name &
                         reg_results$Metric == "RMSE", ]
  bar_vals <- rbind(df_rmse$Original, df_rmse$Anonymized)
  colnames(bar_vals) <- df_rmse$Model

  y_range <- range(bar_vals, na.rm = TRUE)
  y_pad   <- diff(y_range) * 0.2
  ylim    <- c(0, y_range[2] + y_pad)

  barplot(bar_vals, beside = TRUE, col = c("steelblue", "coral"),
          main = paste(ds_name, "- RMSE"), ylab = "RMSE", ylim = ylim,
          names.arg = df_rmse$Model, las = 2,
          legend.text = c("Original", "Anonymized"),
          args.legend = list(x = "topright", cex = 0.7))

  # R-squared
  df_r2 <- reg_results[reg_results$Dataset == ds_name &
                       reg_results$Metric == "R2", ]
  bar_vals <- rbind(df_r2$Original, df_r2$Anonymized)
  colnames(bar_vals) <- df_r2$Model

  y_range <- range(bar_vals, na.rm = TRUE)
  y_pad   <- diff(y_range) * 0.2
  ylim    <- c(min(0, y_range[1] - y_pad), max(1, y_range[2] + y_pad))

  barplot(bar_vals, beside = TRUE, col = c("steelblue", "coral"),
          main = paste(ds_name, "- R-squared"), ylab = "R²", ylim = ylim,
          names.arg = df_r2$Model, las = 2,
          legend.text = c("Original", "Anonymized"),
          args.legend = list(x = "topright", cex = 0.7))
}
```

## Scaling Benchmark

How does anonymization time scale with dataset size and dimensionality?

```{r scaling, message=FALSE, warning=FALSE}
# scale_rows <- c(500, 1000, 2000, 5000, 10000, 20000, 50000)
# scale_cols <- c(10, 25, 50, 100, 200)
# 
# scale_results <- list()
# idx <- 1
# 
# for (p in scale_cols) {
#   for (n in scale_rows) {
#     X <- as.data.frame(matrix(rnorm(n * p), nrow = n))
#     names(X) <- paste0("V", seq_len(p))
# 
#     t0 <- proc.time()
#     anonymize_features(X, key)
#     elapsed <- (proc.time() - t0)[["elapsed"]]
# 
#     scale_results[[idx]] <- data.frame(
#       Rows = n, Features = p, Time_Sec = round(elapsed, 3),
#       stringsAsFactors = FALSE
#     )
#     idx <- idx + 1
#   }
# }
# 
# scale_df <- do.call(rbind, scale_results)
# knitr::kable(scale_df, caption = "Anonymization Time: Rows x Features")
```

```{r scaling_plot, fig.width=10, fig.height=5}
# par(mfrow = c(1, 2), mar = c(5, 4, 3, 1))
# 
# # Time vs rows for each column count
# cols_list <- unique(scale_df$Features)
# colors <- c("steelblue", "coral", "seagreen")
# 
# plot(NULL, xlim = range(scale_df$Rows), ylim = range(scale_df$Time_Sec),
#      xlab = "Rows", ylab = "Time (sec)", main = "Anonymization Time vs Rows")
# for (i in seq_along(cols_list)) {
#   sub <- scale_df[scale_df$Features == cols_list[i], ]
#   lines(sub$Rows, sub$Time_Sec, col = colors[i], lwd = 2, type = "b", pch = 16)
# }
# legend("topleft", legend = paste(cols_list, "features"),
#        col = colors, lwd = 2, pch = 16, cex = 0.8)
# 
# # Time vs features for each row count
# rows_list <- unique(scale_df$Rows)
# colors2 <- hcl.colors(length(rows_list), palette = "Dark 2")
# 
# plot(NULL, xlim = range(scale_df$Features), ylim = range(scale_df$Time_Sec),
#      xlab = "Features", ylab = "Time (sec)",
#      main = "Anonymization Time vs Features")
# for (i in seq_along(rows_list)) {
#   sub <- scale_df[scale_df$Rows == rows_list[i], ]
#   lines(sub$Features, sub$Time_Sec, col = colors2[i], lwd = 2, type = "b", pch = 16)
# }
# legend("topleft", legend = paste(rows_list, "rows"),
#        col = colors2, lwd = 2, pch = 16, cex = 0.8)
```

## Per-Dataset Summary

```{r summary}
cat("Datasets:", length(dataset_names), "\n")
cat("Models:   ranger, gbm, glmnet, lda/ridge\n")
cat("Method:   cryptoencoder anonymization\n\n")

for (ds_name in dataset_names) {
  df <- all_results[all_results$Dataset == ds_name, ]
  task <- df$Task[1]
  n_feat <- df$Features[1]
  n_rows <- df$Rows[1]
  cat(sprintf("--- %s (%d rows, %d features, %s) ---\n",
              ds_name, n_rows, n_feat, task))

  if (task == "classification") {
    acc <- df[df$Metric == "Accuracy", ]
    avg_drop <- mean(acc$Original - acc$Anonymized, na.rm = TRUE)
    cat(sprintf("  Avg accuracy change: %+.4f\n", -avg_drop))

    f1 <- df[df$Metric == "F1" & !is.na(df$Original), ]
    if (nrow(f1) > 0) {
      avg_f1_drop <- mean(f1$Original - f1$Anonymized, na.rm = TRUE)
      cat(sprintf("  Avg F1 change:       %+.4f\n", -avg_f1_drop))
    }
  } else {
    rmse <- df[df$Metric == "RMSE", ]
    avg_rmse_change <- mean(rmse$Anonymized - rmse$Original, na.rm = TRUE)
    cat(sprintf("  Avg RMSE change: %+.4f\n", avg_rmse_change))

    r2 <- df[df$Metric == "R2", ]
    avg_r2_change <- mean(r2$Anonymized - r2$Original, na.rm = TRUE)
    cat(sprintf("  Avg R² change:   %+.4f\n", avg_r2_change))
  }
  cat("\n")
}
```
