---
title: "ML-Safe Anonymization with safetrain"
author: "safetrain"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ML-Safe Anonymization}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(safetrain)
```

## Overview

**safetrain** provides reversible, key-based anonymization of tabular data.
All columns are encoded to numeric and transformed jointly via a
key-derived orthogonal rotation. The rotation matrix and centering/scaling
parameters are derived from the secret key and are **not** stored in the
mapping, so reversal is only possible with the correct key.

When two or more columns are anonymized, the last rotated column is dropped
and encrypted, producing an output data frame with one fewer column. This
means the anonymized data cannot be trivially mapped back to the original
even if the mapping is leaked -- the key is required.

## Quick start

```{r quick}
# Generate a secret key (or use a passphrase with key_from_passphrase())
key <- generate_key()

# Anonymize a data frame
data(sample_data)
res <- anonymize_data(sample_data, key)
anon <- res$data
mapping <- res$mapping

# Anonymized output has one fewer column, with generic names
cat("Original columns:", ncol(sample_data), "\n")
cat("Anonymized columns:", ncol(anon), "\n")
head(anon)
```

## Reversing anonymization

Only with the same key and the mapping object can you recover the original data:

```{r reverse}
back <- deanonymize_data(anon, key, mapping)
# Round-trip recovers originals (within floating-point tolerance for numerics)
all(back$user_id == sample_data$user_id)
all(back$region == sample_data$region)
all(abs(back$amount - sample_data$amount) < 1e-8)
```

## Anonymizing a subset of columns

```{r subset}
res2 <- anonymize_data(sample_data, key, columns = c("user_id", "region"))
# Only user_id and region are anonymized; others unchanged
identical(res2$data$product, sample_data$product)
```

## Model performance on deanonymized data

Because deanonymization exactly recovers the original data, a model trained
on deanonymized data achieves identical performance to one trained on the
original. This verifies the round-trip fidelity.

```{r model-perf, eval = requireNamespace("ranger", quietly = TRUE)}
if (requireNamespace("ranger", quietly = TRUE)) {
  set.seed(42)
  n <- nrow(sample_data)
  idx <- sample(n, floor(0.7 * n))
  train_orig <- sample_data[idx, ]
  test_orig <- sample_data[-idx, ]

  key <- generate_key()
  anon <- anonymize_data(train_orig, key)
  train_recovered <- deanonymize_data(anon$data, key, anon$mapping)

  m_orig <- ranger::ranger(amount ~ region + product + count, data = train_orig)
  m_recovered <- ranger::ranger(amount ~ region + product + count, data = train_recovered)

  pred_orig <- predict(m_orig, test_orig)$predictions
  pred_recovered <- predict(m_recovered, test_orig)$predictions

  rmse_orig <- sqrt(mean((pred_orig - test_orig$amount)^2))
  rmse_recovered <- sqrt(mean((pred_recovered - test_orig$amount)^2))

  cat("RMSE (original): ", round(rmse_orig, 6), "\n")
  cat("RMSE (deanonymized): ", round(rmse_recovered, 6), "\n")
  cat("Difference: ", abs(rmse_orig - rmse_recovered), "\n")
}
```

```{r model-skip, eval = !requireNamespace("ranger", quietly = TRUE)}
# Install the ranger package to run the model performance comparison.
```

## Summary

- Use `generate_key()` or `key_from_passphrase()` to get a secret key.
- Use `anonymize_data(data, key)` to get anonymized data and a mapping; store both the key and the mapping if you need to reverse later.
- Use `deanonymize_data(anon_data, key, mapping)` to recover the original data.
- The anonymization is cryptographically key-dependent: without the key, the rotation cannot be re-derived and the dropped column cannot be decrypted.
