---
title: "ML-Safe Anonymization with safetrain"
author: "safetrain"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ML-Safe Anonymization}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(safetrain)
```

## Overview

**safetrain** provides reversible, key-based anonymization of tabular data.
All columns are encoded to numeric and transformed jointly via a
key-derived orthogonal rotation. The rotation matrix and centering/scaling
parameters are derived from the secret key and are **not** stored in the
mapping, so reversal is only possible with the correct key.

When two or more columns are anonymized, the last rotated column is dropped
and encrypted, producing an output data frame with one fewer column. This
means the anonymized data cannot be trivially mapped back to the original
even if the mapping is leaked -- the key is required.

## Quick start

```{r quick}
# Generate a secret key (or use a passphrase with key_from_passphrase())
key <- generate_key()

# Anonymize a data frame
data(sample_data)
res <- anonymize_data(sample_data, key)
anon <- res$data
mapping <- res$mapping

# Anonymized output has one fewer column, with generic names
cat("Original columns:", ncol(sample_data), "\n")
cat("Anonymized columns:", ncol(anon), "\n")
head(anon)
```

## Reversing anonymization

Only with the same key and the mapping object can you recover the original data:

```{r reverse}
back <- deanonymize_data(anon, key, mapping)
# Round-trip recovers originals (within floating-point tolerance for numerics)
all(back$user_id == sample_data$user_id)
all(back$region == sample_data$region)
all(abs(back$amount - sample_data$amount) < 1e-8)
```

## Anonymizing a subset of columns

```{r subset}
res2 <- anonymize_data(sample_data, key, columns = c("user_id", "region"))
# Only user_id and region are anonymized; others unchanged
identical(res2$data$product, sample_data$product)
```

## Model performance on anonymized data

The cryptoencoder method allows models to be trained directly on the
anonymized representation. The encoder is fit on the training split only;
the same trained weights (stored encrypted in the mapping) are then applied
to the test split via `transform_new_data()`, so both splits share the same
representation space without any leakage.

```{r model-perf, eval = requireNamespace("ranger", quietly = TRUE)}
if (requireNamespace("ranger", quietly = TRUE)) {
  set.seed(42)
  n   <- nrow(sample_data)
  idx <- sample(n, floor(0.7 * n))

  # Feature/target split (target is never anonymized)
  features <- c("region", "product", "count")
  target   <- "amount"

  X_train <- sample_data[idx,  features, drop = FALSE]
  X_test  <- sample_data[-idx, features, drop = FALSE]
  y_train <- sample_data[idx,  target]
  y_test  <- sample_data[-idx, target]

  key <- generate_key()

  # --- Original ---
  train_orig        <- X_train; train_orig[[target]] <- y_train
  test_orig         <- X_test;  test_orig[[target]]  <- y_test
  m_orig            <- ranger::ranger(amount ~ ., data = train_orig)
  pred_orig         <- predict(m_orig, test_orig)$predictions
  rmse_orig         <- sqrt(mean((pred_orig - y_test)^2))

  # --- Anonymized (cryptoencoder): fit on train, apply to test ---
  res_train         <- anonymize_data(X_train, key,
                                     opts = list(method = "cryptoencoder"))
  anon_train        <- res_train$data; anon_train[[target]] <- y_train

  anon_test         <- transform_new_data(X_test, key, res_train$mapping)
  anon_test[[target]] <- y_test

  m_anon            <- ranger::ranger(amount ~ ., data = anon_train)
  pred_anon         <- predict(m_anon, anon_test)$predictions
  rmse_anon         <- sqrt(mean((pred_anon - y_test)^2))

  cat("RMSE (original features):   ", round(rmse_orig, 4), "\n")
  cat("RMSE (anonymized features): ", round(rmse_anon, 4), "\n")
  cat("Difference:                 ", round(abs(rmse_orig - rmse_anon), 4), "\n")
}
```

```{r model-skip, eval = !requireNamespace("ranger", quietly = TRUE)}
# Install the ranger package to run the model performance comparison.
```

## Differential privacy

The cryptoencoder output is bounded in (−1, 1) per column due to tanh. This
makes Gaussian noise calibration unusually clean: the global L2 sensitivity of
any single row is `2 * sqrt(n)`, where n is the number of columns. Pass
`dp_epsilon` (and optionally `dp_delta`, default `1e-5`) to add
(ε, δ)-differentially private noise after the post-rotation step.

```{r dp-example}
key_dp <- generate_key()
data(sample_data)
X <- sample_data[, c("amount", "count", "user_id")]

res_dp <- anonymize_data(X, key_dp,
                         opts = list(method = "cryptoencoder",
                                     dp_epsilon = 1.0,
                                     dp_delta   = 1e-5))

cat("Noise sigma applied: ", round(res_dp$mapping$pca$dp_sigma, 4), "\n")
cat("DP guarantee:        (epsilon =", res_dp$mapping$pca$dp_epsilon,
    ", delta =", res_dp$mapping$pca$dp_delta, ")\n")
```

Smaller `dp_epsilon` means stronger privacy and more noise. A common choice for
`dp_delta` is `1 / nrow(data)`. The noise level is stored in the mapping for
audit purposes.

Note: the (ε, δ)-DP guarantee applies to the **encoding step with fixed
pre-trained weights** — i.e. when using `transform_new_data()` to encode new
individuals through an already-trained encoder. It does not cover the
autoencoder training phase itself. Organisations requiring full-pipeline
differential privacy should apply DP-SGD during training.

## Key destruction and GDPR anonymization

By default, `anonymize_data()` returns a mapping that allows full recovery with
the key — this is **pseudonymization** under GDPR Article 4(5), and the data
remains personal data. Set `destroy_key = TRUE` when auditability is not
required: the key is zeroed inside the function and the mapping is permanently
locked, converting the scheme to **de facto anonymization** under GDPR
Recital 26.

```{r destroy-key}
key_anon <- generate_key()
data(sample_data)

res_anon <- anonymize_data(sample_data, key_anon, destroy_key = TRUE)

# The mapping records that destruction occurred
cat("Key destroyed: ", res_anon$mapping$key_destroyed, "\n")

# Attempting recovery now errors immediately
tryCatch(
  deanonymize_data(res_anon$data, key_anon, res_anon$mapping),
  error = function(e) cat("Recovery blocked:", conditionMessage(e), "\n")
)
```

The table below summarises the GDPR compliance modes:

| `destroy_key` | `dp_epsilon` | GDPR status | Auditability |
|---|---|---|---|
| `FALSE` (default) | `NULL` | Pseudonymization (Art. 4(5)) | Full recovery with key |
| `FALSE` | set | Pseudonymization + formal (ε,δ)-DP | Full recovery (noisy) |
| `TRUE` | `NULL` | De facto anonymization (Recital 26) | None |
| `TRUE` | set | De facto anonymization + (ε,δ)-DP | None |

## Summary

- Use `generate_key()` or `key_from_passphrase()` to get a secret key.
- Use `anonymize_data(data, key)` to get anonymized data and a mapping; store
  both the key and the mapping if you need to reverse later.
- Use `deanonymize_data(anon_data, key, mapping)` to recover the original data.
- Use `transform_new_data(new_data, key, mapping)` to encode new observations
  through an already-trained cryptoencoder without retraining.
- Set `opts$dp_epsilon` to add (ε, δ)-differential privacy noise
  (cryptoencoder only).
- Set `destroy_key = TRUE` to permanently prevent recovery, converting
  pseudonymization to de facto anonymization under GDPR Recital 26.
